---


title: "IVMOOC - Global Trade Visualisation Tool"
author: "Ronny De Winter"
date: "18 Mar 2015"
output: html_document
---

This document descibes some explorations on the global trade data (data set tstrade.csv) as provided by the University on Indiana, for a Client Project "Global Trade Visualization Tool" of the Purdue University.

Load the data:

```{r load, cache=TRUE}
tstrade <- read.csv("../tstrade.csv")
```

Obtain basic data information:

```{r explore}
str(tstrade)
unique(tstrade$YEAR)
```

There are 15 million observations of 5 variables, data on 57 trading commodities, 134 regions, from 1995 to 2009.
This represents a full matrix across the 4 dimensions: 57 * 134 * 134 * 15 = 15352380

**Question:** There are 2 REG variables, are these the exporting and importing regions (in that order)?  
**Question:** What is the definition of the Value variable? export of TRAD_COMM from REG in USD, mUSD?

```{r tstradeSummary, cache=TRUE}
summary(tstrade)
```

There are 74 negative "Value" values, REG related:
```{r negVal-REG}
negVal = subset(tstrade, tstrade$Value < 0)
subset(table(negVal$REG),table(negVal$REG)>0)
```
REG.1 related:
```{r negVal-REG.1}
subset(table(negVal$REG.1),table(negVal$REG.1)>0)
```
**Question:** Is this related to indirect (re-) trades? Do we have to include them in the visualizations?

Distribution of the "Value" variable

```{r ValueHist}
hist(log(tstrade$Value))
```

2 main groups can be idenfied: very small values (< 1E-5) and larger values (> 1E-2).
Let's assume that the very small values are placeholders for zero or not available (NA) data and remove these form the dataset. We also ignore the negative values from now.

```{r removeNA, cache=TRUE}
# The theeshold of 3 is possible to be increased to get a smaller more relevant dataset 
tstrade2 = subset(tstrade, tstrade$Value > 3)
nrow(tstrade2)
summary(tstrade2$Value)
boxplot(tstrade2$Value)
hist(log(log(tstrade2$Value)))
```
